\textbf{Predicting virus-receptor mutant binding by molecular dynamics simulation}

### 2013-07-08 paper.tex
Slight change in the mutant comparison figure.
I wanted to push changes before deleting more text.

Existing computational methods to predict protein--protein interaction affinity often perform poorly in important test cases. In particular, the effects of multiple mutations, non-alanine substitutions, and flexible loops are difficult to predict with available tools and protocols. We present here a new method to interrogate affinity differences resulting from mutations in a host-virus protein--protein interface. Our method is based on extensive non-equilibrium all-atom simulations: We computationally pull the machupo virus (MACV) spike glycoprotein (GP1) away from the human transferrin receptor (hTfR1) and estimate affinity using the maximum applied force during a pulling simulation and the area under the force-versus-distance curve. We find that these quantities can provide novel biophysical insight into the GP1/hTfR1 interaction. First, with no prior knowledge of the system we can differentiate among wild-type and mutant complexes. Second, although the static co-crystal structure shows two large hydrogen-bonding networks in the GP1/hTfR1 interface, our simulations indicate that only one of them is critical for the binding interaction. Third, one viral site known to be critical for infection may mark an important evolutionary suppressor site for infection-resistant hTfR1 mutants. Finally, our method provides an elegant framework to compare the effects of multiple mutations, individually and jointly, on protein--protein interactions.
\end{abstract}

@@

We applied our computational method to wild type (WT) and mutant complexes, and found that we could resolve relative differences in unbinding and predict significant affinity changes. At sites known to be important for successful viral entry, we found that the biochemical cause of reduced infectivity may not be as simple as the static structure suggests. For example, the static structure shows a hydrogen-bonding network connected to site Asn348 in hTfR1. According to our simulations, this network likely does not affect binding affinity directly. Our study offers an all-atom steered molecular dynamic approach to avoid the pitfalls of several existing methods used to evaluate mutations in protein--protein interfaces.

@@

Finally, SMD was run for 15 ns (7,500,000 time steps) of chemical time. For each simulation, we randomly selected one of the equilibration runs for restart. We ran 50 replicate simulations per mutant. All GP1/hTfR1 complexes separated by greater than 4 \AA\ and many separated to 10 or more.

@@

The statistical package R was used for all further analysis and visualization. Each of the 50 independent trajectories per mutant produced a fairly noisy force curve. To reduce the noise, we averaged each point with a 75 point interpolation window. The window ensured that the averaged maximum applied force was never cut off near the beginning of the simulation due to lost time steps (caused by averaging). The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC). We used the pairwise.t.test function with the FDR p-value correction to test for significance. The ggplot \citep{ggplot} package was used to generate most of the figures.

@@

In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants}). Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}. Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 or Ser111 in GP1. The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \citep{Abraham2010}. In a manner similar to site 211, Asn348 appears to be important for binding by participating in a critical hydrogen bonding network \citep{Rad2008,Abraham2010} to GP1. In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}. Finally, an alanine mutation at site 111 in GP1 has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}. For notation purposes, the viral site is always referred to with a preceding `v'.

@@

We found that relative to WT, one alanine mutation (Tyr211Ala) produced a very large and statistically significant difference in the maximum applied force and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:pairwise_differences}), while the other two did not (Table~\ref{tab:pairwise_differences}). From these results, we concluded that our method can indeed successfully distinguish individual mutants from WT, and that alanine substitutions at Asn348 and vArg111 may have little effect on binding affinity (see below for further discussion). When considering additional mutants (also discussed below), we found that maximum applied force was generally sufficient to distinguish mutants (Tables~\ref{tab:pairwise_differences} and~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}). In general, however, maximum applied force seemed to be the more sensitive statistic than AUC.

\subsection*{Comparative analysis of the GP1/hTfR1 interface}
Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface, it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction. One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function. Seen in Table~\ref{tab:summary_mutant}), we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine. Several mutations caused significant relative affinity changes. In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause reduced binding. Then, we compared the size of those differences to reduced binding single mutants (Figure~\ref{fig:sensitivity}).

Although Tyr211Ala appears to have a large impact on binding affinity, no single mutant can provide enough evidence to understand the biochemical difference in binding mechanism. Since alanine is both smaller than tyrosine and also incapable of participating in hydrogen-bond interactions, we tested further mutations to identify the critical biochemical difference responsible for loss of binding affinity. In particular, we substituted smaller side chains that, like tyrosine, were capable of hydrogen bonding. We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}. Both mutations proved incapable of causing reduced binding affinity in our simulations. In fact, they caused a significant increase in relative binding affinity (Figure~\ref{fig:sensitivity} and Table~\ref{tab:MAF_p}).

We also simulated several point mutations at Asn348 in the hTfR1. As discussed above, the alanine mutation at this site showed no significant difference in maximum applied force or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}). In addition, the Asn348Lys mutation showed no significant difference from WT, and neither did Asn348Trp. (For both of these mutations, however, mean maximum applied force and mean AUC was lower than for WT. See Table~\ref{tab:summary_statistics}.)  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder. Moreover, Asn348Trp showed nearly identical results to Asn348Lys. These results suggest that, in contrast to Tyr211, the critical difference at position 348 in hTfR1 is size exclusion rather than the destruction of a hydrogen-bonding network. The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:pairwise_differences}). To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding. As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $) (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study. Further, according to maximum applied force (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}). We suspect that the effect of Asn348Trp alone is near the limit of detection using our method. A larger number of replicates would possibly have resolved affinity differences between Asn348Trp and WT or other mutants more consistently.

Last, we further analyzed a single mutation in GP1, vArg111Ala. As mentioned in the previous subsection, in our simulations this mutant showed no significant change in either maximum applied force or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}). This result was somewhat surprising, since Tyr211Ala, presumably disrupting the same hydrogen-bonding network as vArg111Ala, displayed a significant reduction in affinity. To probe the interaction between position 111 in the GP1 and position 211 in the hTfR1 further, we also tested the double mutant vArg111Ala/Tyr211Ala. This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:pairwise_differences}). Thus, vArg111Ala in the GP1 appeared capable of suppressing the loss-of-binding caused by Tyr211Ala. This result shows that the two sites do indeed interact significantly, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction could lead to comparable binding affinity.

@@

We have applied a method utilizing steering forces in all-atom molecular dynamics simulations to evaluate the effects of mutations at the GP1/hTfR1 interface. We modeled mutations at several sites in the GP1/hTfR1 interface, and verified that our computational protocol was sensitive enough to distinguish point mutants in hTfR1. Further, we identified two test statistics, maximum applied force and AUC, that can be used as proxies for binding affinity. We systematically tested several point mutations to understand their contribution to the binding interaction. In the case of Asn348Lys, we have shown that the static structure provides little insight into why this mutation causes loss-of-infectivity \textit{in vivo}. While Asn348 appears to be involved in a hydrogen-bonding network in the static structure, loss of binding affinity at that site is actually caused by size and charge restriction. Further, we have shown that the cause of diminished infectivity in the viral single-mutant vArg111Ala may not be due to binding-energy differences. We also found that a negatively polar residue at site 211 in hTfR1 is absolutely critical for a tight binding interaction. Any non-polar mutation at Tyr211 in hTfR1 is likely to completely halt viral entry and dramatically decrease the chances of MACV infection. However, our data suggest that the viral mutation vArg111Ala may restore the infective potential of MACV that was previously destroyed by a Tyr211Ala hTfR1 mutation.

By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application. Traditionally this technique is either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Buch2011,Giorgino2012}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}. Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2012}. And still others aimed at more esoteric problems, such as applying driving forces to untie polypeptide knots in toy systems \citep{Sulkowska2010}. Here, we have shown that SMD can provide insight into the \textit{relative} strength of protein--protein interactions. In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application. Further, our method complements whole cell infectivity assays, which demonstrate phenotypic effects but do not provide biochemical mechanisms. Via SMD, one can separate mutations whose likely effect is altered binding affinity from mutations that may affect other factors such as protein expression, protein folding, or interactions with other proteins not considered in the simulation. Thus, SMD may open avenues for subsequent experimental work.

Our findings may rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}. First, we found that some substitutions at positions 211 and 348 did affect the strength of receptor binding. However, the computational data suggest that the reason and nature of the effects at these two sites are very different. Congruent with entry data \citep{Rad2008,Rad20111}, mutations at Tyr211 often have a dramatic effect on receptor binding. Second, no individual point mutation at Asn348 showed a significantly different affinity from the WT complex. However, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference. That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually. Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations. Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone. As mutations at hTfR1 site 211 are known to reduce infectivity \emph{in vitro} and \emph{in vivo}, it is possible that vArg111 in GP1 can serve as a site for compensatory MACV mutations.

@@

There are a few additional challenges for investigating host-virus interaction via molecular dynamics simulation. As with any atomistic simulation, there is going to be a fairly large noise to signal ratio. To reduce noise, one could further customize each simulation, e.g. by determining the optimal pulling speed. Furthermore, larger amounts of computational resources will have a direct and powerful impact on the strength of any atomistic study \citep{Shaw2012}. Such resources could come in the form of increased compute time, improved code, or customized hardware for floating point operations \citep{Shaw2011}. With improved resources, we could investigate thousands of individual permutations in the GP1/hTfR1 binding interface. Going forward, we believe that all-atom molecular dynamics simulations can become an increasingly viable approach for protein interaction studies.

@@

\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=6in]{figures/sensitivity_histogram.pdf}}
\caption{\label{fig:sensitivity}\title{Distribution of interpolated maximum force (MAF) for three different GP1/hTfR1 complexes.} The WT GP1-hTfR1 complex in the middle is flanked by the tighter binding mutant Y211D on the right and the weaker binding double mutant N348W/Y211A on the left. The large non-overlapping areas indicate a large and statistically significant difference in these three complexes.}
\end{figure}

\clearpage
\begin {table}[H]
\caption{\label{tab:summary_mutant}Summary of information for each mutation tested. Observed $in~vivo$ refers to mutations that have been observed in rodent populations. Phenotype $in~vitro$ refers to the observed phenotype in $in~vitro$ viral entry assays.}
%Phenotype $in~silico$ refers to mutations that were statistically significant relative to wild-type only. 
\begin{center}
  \resizebox{11cm}{!} {
    \begin{tabular}{l c r r}
    \hline
      Mutation & Observed $in~vivo$ & Phenotype $in~vitro$ \\%& Phenotype $in~silico$ \\ \hline
            WT &                Yes &         Normal Entry \\%&        Normal Binding \\
         N348A &                 No &                    - \\%&       Not Significant \\
         N348K &                Yes &     Diminished Entry \\%&       Not Significant \\
         N348W &                 No &                    - \\%&       Not Significant \\
         Y211A &                 No &        No Expression \\%&     Decreased Binding \\ 
         Y211D &                Yes &        No Expression \\%&     Increased Binding \\ 
         Y211T &                 No &     Diminished Entry \\%&     Increased Binding \\
        vR111A &                 No &     Diminished Entry \\%&       Not Significant \\
   N348A/Y211A &                 No &                    - \\%&       Not Significant \\
   N348W/Y211A &                 No &                    - \\%&     Decreased Binding \\
  vR111A/Y211A &                 No &                    - \\%&       Not Significant \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

@@

\clearpage
\begin {table}[H]
\caption{\label{tab:pairwise_differences}Pairwise differences (row variable minus column variable) in mean maximum applied force. Bolded values are statistically significant at $p<0.05$. Values marked with * and ** are significant at $p<0.01$ and $p<0.001$, respectively.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
                & WT                  & N348A              & N348W              & N348K              & Y211A             & Y211D              & Y211T              & vR111A             & N348A/Y211A        & N348W/Y211A        \\ \hline
    N348A       &  27.33              &                    &                    &                    &                   &                    &                    &                    &                    &                    \\
    N348W       & -39.87              & \textbf{-67.20}    &                    &                    &                   &                    &                    &                    &                    &                    \\
    N348K       & -40.64              & \textbf{-67.97}    & -0.77              &                    &                   &                    &                    &                    &                    &                    \\
    Y211A       & \textbf{-87.97*}    & \textbf{-115.30**} & -48.10             & -47.33             &                   &                    &                    &                    &                    &                    \\
    Y211D       & \textbf{103.44**}   & \textbf{76.12*}    & \textbf{143.31**}  & \textbf{144.08**}  & \textbf{191.42**} &                    &                    &                    &                    &                    \\
    Y211T       & \textbf{89.17*}     & \textbf{61.85}     & \textbf{129.04**}  & \textbf{129.81**}  & \textbf{177.15**} & -14.27             &                    &                    &                    &                    \\
   vR111A       & -17.76              & -44.94             & 22.26              & 23.03              & \textbf{70.36}    & \textbf{-121.05**} & \textbf{-106.78**} &                    &                    &                    \\
    N348A/Y211A & -9.04               & -36.37             & 30.83              & 31.60              & \textbf{78.93}    & \textbf{-112.48**} & \textbf{-98.21*}   & 8.57               &                    &                    \\
    N348W/Y211A & \textbf{-154.79**}  & \textbf{-182.12**} & \textbf{-114.92**} & \textbf{-114.15**} & \textbf{-66.2}    & \textbf{-258.23**} & \textbf{-243.96**} & \textbf{-137.18**} & \textbf{-145.75**} &                    \\
   vR111A/Y211A &  8.71               & -18.12             & 48.58              & 49.35              & \textbf{96.68*}   & \textbf{-94.74*}   & \textbf{-80.46*}   & 26.31              & 0.59               & \textbf{163.50**} \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

>>>>

Existing computational methods to predict protein--protein interaction affinity often perform poorly in important test cases. In particular, the effects of multiple mutations, non-alanine substitutions, and flexible loops are difficult to predict with available tools and protocols. We present here a new method to interrogate affinity differences resulting from mutations in a host-virus protein--protein interface. Our method is based on extensive non-equilibrium all-atom simulations: We computationally pull the machupo virus (MACV) spike glycoprotein (GP1) away from the human transferrin receptor (hTfR1) and estimate affinity using the maximum applied force during a pulling simulation and the area under the force-versus-distance curve. We find that these quantities can provide novel biophysical insight into the GP1/hTfR1 interaction. First, with no prior knowledge of the system we can differentiate among wild-type and mutant complexes. Second, although the static co-crystal structure shows two large hydrogen-bonding networks in the GP1/hTfR1 interface, our simulations indicate that one of them may not be important for tight binding. Third, one viral site known to be critical for infection may mark an important evolutionary suppressor site for infection-resistant hTfR1 mutants. Finally, our method provides an elegant framework to compare the effects of multiple mutations, individually and jointly, on protein--protein interactions.
\end{abstract}

@@

We applied our computational method to wild type (WT) and mutant complexes, and found that we could resolve relative differences in unbinding and predict significant affinity changes. At sites known to be important for successful viral entry, we found that the biochemical cause of reduced infectivity may not be as simple as the static structure suggests. For example, the static structure shows a hydrogen-bonding network connected to site Asn348 in hTfR1. According to our simulations, this network may not affect binding affinity directly. In addition, our study offers an all-atom steered molecular dynamic approach to avoid the pitfalls of several existing methods used to evaluate mutations in protein--protein interfaces.

@@

Finally, SMD was run for 15 ns (7,500,000 time steps) of chemical time. For each simulation, we randomly selected one of the equilibration runs for restart. We ran 50 replicate simulations per mutant for a total of 550 SMD simulations. All GP1/hTfR1 complexes separated by greater than 4 \AA\ and many separated to 10 or more.

@@

The statistical package R was used for all further analysis and visualization. Each of the 50 independent trajectories per mutant produced a fairly noisy force curve. To reduce the noise, we averaged each point with a 50 point interpolation window. The window ensured that the averaged maximum applied force was never cut off near the beginning of the simulation due to lost time steps (caused by averaging). The two primary descriptive statistics we used were maximum interpolated applied force and total area under the interpolated curve (AUC). We used the pairwise.t.test function with the FDR p-value correction to test for significance. The ggplot \citep{ggplot} package was used to generate most of the figures.

@@

In total, we tested 7 point mutants and 3 double mutants in addition to the WT complex (Table~\ref{tab:summary_mutants} and Figure~\ref{fig:mutant_comparison}). Mutations at site hTfR1 site 211 have proven capable of causing loss-of-entry according to \textit{in vitro} flow-cytometry infection assays or known host-range limitations \citep{Rad2008,Rad20111,Rad20112}. Most likely, this effect is caused by the destruction of a critical hydrogen bond to Ser113 or Ser111 in GP1. The lost hydrogen bond would lead to the subsequent loss of a large hydrogen-bonding network seen in the crystal structure (Table~\ref{tab:summary_mutants}) \citep{Abraham2010}. In a manner similar to site 211, Asn348 appears to be important for binding by participating in a critical hydrogen bonding network \citep{Rad2008,Abraham2010} to GP1. In particular, Asn348Lys is reported in the literature to cause significantly reduced viral entry \textit{in vivo} (Table~\ref{tab:summary_mutants}) \citep{Rad2008,Abraham2010}. Finally, an alanine mutation at site 111 in GP1 has also been shown to cause decreased entry (Table~\ref{tab:summary_mutants}) \citep{Rad20112}. For notation purposes, the viral site is always referred to with a preceding `v'.

@@

We found that relative to WT, one alanine mutation (Tyr211Ala) produced a very large and statistically significant difference in the maximum applied force and AUC (Figure~\ref{fig:force_curve}, Table~\ref{tab:MAF_pairwise}), while the other two did not (Table~\ref{tab:MAF_pairwise}). When considering additional mutants (also discussed below), we found that maximum applied force was generally sufficient to distinguish mutants (Tables~\ref{tab:MAF_pairwise} and ~\ref{tab:MAF_p}), and AUC was able to add a few more statistically significant differences (Table~\ref{tab:AUC_p}). In general, however, maximum applied force seemed to be the more sensitive statistic than AUC.

\subsection*{Comparative analysis of the GP1/hTfR1 interface}
Considering the involvement of extended hydrogen-bonding networks in the GP1/hTfR1 interface, it was not clear that individual alanine mutations, even those that should destroy such networks, would significantly change the strength of interaction. One major advantage of first principles simulations is the ability to test mutations other than alanine without additional underlying assumptions in the energy function. Seen in Table~\ref{tab:summary_mutant}), we made additional mutations based on biochemical intuition or available experimental data to chemically diverse amino acids including tryptophan, lysine, aspartate, and threonine. Several mutations caused significant relative affinity changes. In addition, to detect synergistic effects, we tested several double mutants where both mutations appeared to cause similar changes in binding. Then, we compared the size of those differences to single mutants (Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}).

Although Tyr211Ala appears to have a large impact on binding affinity, no single mutant can provide enough evidence to understand the biochemical difference in binding mechanism. Since alanine is both smaller than tyrosine and also incapable of participating in hydrogen-bond interactions, we tested further mutations to identify the critical biochemical difference responsible for change in binding affinity. In particular, we substituted smaller side chains that, like tyrosine, were capable of hydrogen bonding. We chose Tyr211Asp and Tyr211Thr, two mutations that have been discussed in the context of selection pressure on hosts in rodent populations \citep{Rad2008,Rad20111,Rad20112}. Both mutations proved capable of causing a significant change in binding affinity in our simulations, but the change appeared to be increased affinity. (Figure~\ref{fig:sensitivity} and ~\ref{fig:mutant_comparison}, and Table~\ref{tab:MAF_p}).

We also simulated several point mutations at Asn348 in the hTfR1. As discussed above, the alanine mutation at this site showed no significant difference in maximum applied force or AUC from WT (Tables~\ref{tab:MAF_p} and ~\ref{tab:AUC_p}). In addition, the Asn348Lys mutation showed no significant difference from WT, and neither did Asn348Trp. (For both of these mutations, however, mean maximum applied force and mean AUC was lower than for WT. See Table~\ref{tab:summary_statistics}.)  On the other hand, there was a detectable difference between Asn348Ala and Asn348Lys (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), with Asn348Lys being a weaker binder. Moreover, Asn348Trp showed nearly identical results to Asn348Lys. The mutations to large amino acids (Asn348Trp and Asn348Lys) produced nearly identical affinity changes, whereas the mutations to amino acids not capable of hydrogen bonding (Asn348Ala and Asn348Trp) produced significantly different affinity changes (Table~\ref{tab:MAF_pairwise}). To check the consistency of our results, we hypothesized that the combination of Tyr211Ala and Asn348Trp, being chemically disconnected in two different hydrogen-bonding networks, would lead to a synergistic loss-of-binding. As expected, the double mutant was the weakest binding mutant tested ($ p < 10^{-6} $) (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}) in this study. Further, according to maximum applied force (but not AUC), the combination of Tyr211Ala and Asn348Trp also showed significantly weaker binding than Tyr211Ala by itself (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}). We suspect that the effect of Asn348Trp alone is near the limit of detection using our method. A larger number of replicates would possibly have resolved affinity differences between Asn348Trp and WT or other mutants more consistently.

Last, we further analyzed a single mutation in GP1, vArg111Ala. As mentioned in the previous subsection, in our simulations this mutant showed no significant change in either maximum applied force or AUC (Tables~\ref{tab:MAF_p} and~\ref{tab:AUC_p}), even though both quantities were, on average, lower than in WT (Table~\ref{tab:summary_statistics}). This result was somewhat surprising, since Tyr211Ala, presumably disrupting the same hydrogen-bonding network as vArg111Ala, displayed a significant reduction in affinity. To probe the interaction between position 111 in the GP1 and position 211 in the hTfR1 further, we also tested the double mutant vArg111Ala/Tyr211Ala. This double mutant showed affinity indistinguishable from WT and significantly higher than Tyr211Ala alone  (Table~\ref{tab:MAF_pairwises}). This result shows that the two sites do indeed interact, and that replacing the hydrogen-bonding network at these sites with a hydrophobic interaction could lead to comparable binding affinity.

@@

We have applied a method utilizing steering forces in all-atom molecular dynamics simulations to evaluate the effects of mutations at the GP1/hTfR1 interface. We modeled mutations at several sites in the GP1/hTfR1 interface, and verified that our computational protocol was sensitive enough to distinguish point mutants in hTfR1. Further, we identified two test statistics, maximum applied force and AUC, that can be used as proxies for binding affinity. We systematically tested several point mutations to understand their contribution to the binding interaction. In the case of Asn348Lys, we have shown that the static structure provides little insight into why this mutation causes loss-of-infectivity \textit{in vivo}. While Asn348 appears to be involved in a hydrogen-bonding network in the static structure, change in binding at that site may actually be caused by size and charge restriction. We also found that a negatively polar residue at site 211 in hTfR1 is absolutely critical for a tight binding interaction. Any non-polar mutation at Tyr211 in hTfR1 is likely to completely halt viral entry and dramatically decrease the chances of MACV infection.

By extending the SMD technique to investigate a protein--protein interaction we have, in many ways, changed the scope of its application. Traditionally this technique is either applied to compute equilibrium quantities via a non-equilibrium approximation \citep{Park2003,Park2004,Giorgino2011}, used to estimate protein stability through unfolding \citep{Lu1999}, or used to calculate the absolute free energy of small molecule ligand binding \citep{Dixit2001}. Likewise, others have used SMD to understand the process of binding and unbinding at a resolution unmatched by experiment \citep{Cuendet2011,Giorgino2011}. And still others aimed at more esoteric problems, such as applying driving forces to untie polypeptide knots in toy systems \citep{Sulkowska2010}. Here, we have shown that SMD can provide insight into the \textit{relative} strength of protein--protein interactions. In contrast to more traditional absolute affinity calculations, relative binding is much more useful for experimental application. Further, our method complements whole cell infectivity assays, which demonstrate phenotypic effects but do not provide biochemical mechanisms. Via SMD, one can separate mutations whose likely effect is altered binding affinity from mutations that may affect other factors such as protein expression, protein folding, or interactions with other proteins not considered in the simulation. Thus, SMD may open avenues for subsequent experimental work.

Our findings may rationalize several effects observed in both infectivity data and rodent populations \citep{Rad2008,Rad20111}. First, we found that some substitutions at positions 211 and 348 did affect the strength of receptor binding. However, the computational data suggest that the reason and nature of the effects at these two sites are very different. Congruent with entry data \citep{Rad2008,Rad20111}, mutations at Tyr211 often have a dramatic effect on receptor binding. Second, no individual point mutation at Asn348 showed a significantly different affinity from the WT complex. However, when Tyr211Ala was paired with large (Trp) and positively charged (Lys) substitutions at position 348, the result was a larger than expected synergistic difference. That is, the double mutant Tyr211Ala/Asn348Trp caused a much larger decrease in binding than we expected from either mutation individually. Third, the GP1 mutation vArg111Ala causes a loss-of-infection during \textit{in vitro} infectivity assays \citep{Rad20112}, yet it was indistinguishable from the WT complex in our simulations. Although Tyr211Ala was the most disruptive single mutant we tested, vArg111Ala in the GP1 was able to restore mean maximum applied force to WT levels (Table~\ref{tab:summary_statistics}), and to levels significantly higher than observed for Tyr211Ala alone.

@@

There are a few additional challenges for investigating host-virus interaction via molecular dynamics simulation. As with any atomistic simulation, there is going to be a fairly large noise to signal ratio. To reduce noise, one could further customize each simulation, e.g. by determining the optimal pulling speed. Furthermore, larger amounts of computational resources will have a direct and powerful impact on the strength of any atomistic study \citep{Shaw2012}. Such resources could come in the form of increased compute time, improved code, or customized hardware for floating point operations \citep{Shaw2011}. With improved resources, we could investigate thousands of individual permutations in the GP1/hTfR1 binding interface. In addition, with additional compute time it would be possible to incorporate equilibrium sampling approaches \citep{Buch2011} or use brute force equilibrium approaches \citep{Giorgino2012} to improve resolution. Going forward, we believe that all-atom molecular dynamics simulations can become an increasingly viable approach for protein interaction studies.

@@

\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=6in]{figures/sensitivity_histogram.pdf}}
\caption{\label{fig:sensitivity}\title{Distribution of interpolated maximum force for three different GP1/hTfR1 complexes.} The WT GP1-hTfR1 complex in the middle is flanked by the tighter binding mutant Y211D on the right and the weaker binding double mutant N348W/Y211A on the left. The large non-overlapping areas indicate a large and statistically significant difference in these three complexes.}
\end{figure}

\clearpage
\begin{figure}[p]
\centerline{\includegraphics[width=6in]{figures/mutant_comparison_boxplot.pdf}}
\caption{\label{fig:mutant_comparison}\title{Distribution of interpolated maximum force for all complexes tested.} The statistically significant difference relative to the WT complex are indicated by a star above the distribution.}
\end{figure}

\clearpage
\begin {table}[H]
\caption{\label{tab:summary_mutants}Summary of information for each mutation tested. Observed $in~vivo$ refers to mutations that have been observed in rodent populations. Phenotype $in~vitro$ refers to the observed phenotype in $in~vitro$ viral entry assays.} 
\begin{center}
  \resizebox{11cm}{!} {
    \begin{tabular}{l c r r}
    \hline
      Mutation & Observed $in~vivo$ & Phenotype $in~vitro$ \\
            WT &                Yes &         Normal Entry \\
         N348A &                 No &                    - \\
         N348K &                Yes &     Diminished Entry \\
         N348W &                 No &                    - \\
         Y211A &                 No &        No Expression \\
         Y211D &                Yes &        No Expression \\
         Y211T &                 No &     Diminished Entry \\
        vR111A &                 No &     Diminished Entry \\
   N348A/Y211A &                 No &                    - \\
   N348W/Y211A &                 No &                    - \\
  vR111A/Y211A &                 No &                    - \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}

@@

\clearpage
\begin {table}[H]
\caption{\label{tab:MAF_pairwise}Pairwise differences (row variable minus column variable) in mean maximum applied force. Bolded values are statistically significant at $p<0.05$. Values marked with * and ** are significant at $p<0.01$ and $p<0.001$, respectively.}
\begin{center}
  \resizebox{16cm}{!} {
    \begin{tabular}{l l l l l l l l l l l}
    \hline
                & WT                  & N348A              & N348W              & N348K              & Y211A             & Y211D              & Y211T              & vR111A             & N348A/Y211A        & N348W/Y211A        \\ \hline
    N348A       &  27.33              &                    &                    &                    &                   &                    &                    &                    &                    &                    \\
    N348W       & -39.87              & \textbf{-67.20}    &                    &                    &                   &                    &                    &                    &                    &                    \\
    N348K       & -40.64              & \textbf{-67.97}    & -0.77              &                    &                   &                    &                    &                    &                    &                    \\
    Y211A       & \textbf{-87.97*}    & \textbf{-115.30**} & -48.10             & -47.33             &                   &                    &                    &                    &                    &                    \\
    Y211D       & \textbf{103.44**}   & \textbf{76.12*}    & \textbf{143.31**}  & \textbf{144.08**}  & \textbf{191.42**} &                    &                    &                    &                    &                    \\
    Y211T       & \textbf{89.17*}     & \textbf{61.85}     & \textbf{129.04**}  & \textbf{129.81**}  & \textbf{177.15**} & -14.27             &                    &                    &                    &                    \\
   vR111A       & -17.76              & -44.94             & 22.26              & 23.03              & \textbf{70.36}    & \textbf{-121.05**} & \textbf{-106.78**} &                    &                    &                    \\
    N348A/Y211A & -9.04               & -36.37             & 30.83              & 31.60              & \textbf{78.93}    & \textbf{-112.48**} & \textbf{-98.21*}   & 8.57               &                    &                    \\
    N348W/Y211A & \textbf{-154.79**}  & \textbf{-182.12**} & \textbf{-114.92**} & \textbf{-114.15**} & \textbf{-66.2}    & \textbf{-258.23**} & \textbf{-243.96**} & \textbf{-137.18**} & \textbf{-145.75**} &                    \\
   vR111A/Y211A &  8.71               & -18.12             & 48.58              & 49.35              & \textbf{96.68*}   & \textbf{-94.74*}   & \textbf{-80.46*}   & 26.31              & 0.59               & \textbf{163.50**} \\
    \hline
    \end{tabular}
  }
\end{center}
\end{table}
